DP problems of interest

### 2012-09-28 docs/roadmap/problems.tex
more problem instances

@@

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	M_{(i+1,j-1)}+\omega(i,j) & SW\\
	M_{(i+1,j)} & S\\
	M_{(i,j-1)} & W\\
	\max_{i<k<j}M_{(i,k)}+W_{(k+1,j)} & k
\end{array}\right\} = B_{(i,j)} \]
With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.

% ------------------------------------------------------------------------------------------------
%\newpage
%\subsection{Zuker folding}
% XXX: we let this aside for the moment as the recurrence has too many unknown parameters
%Problem: folding a RNA string $S$ over itself.
%
%Alphabets:\ul
%\item Input: $\Sigma(S)=\{A,C,G,U\}$.
%\item Cost matrix: $\Sigma(W)=\Sigma(V)=\Sigma(F)=\{0..n\},n=$
%\item Backtrack matrix: $\Sigma(B)=\{\}$
%\ule
%
%Initialization:\ul
%\item Cost matrices:\ul
%	\item xxx
%\ule
%\item Backtrack matrix: $B$
%\ule
%
%Recurrence:
%\[\begin{array}{rcl}
%W_{(i,j)}&=&\min\left\{\begin{array}{l}
%	W_{(i+1,j)}+b \\
%	W_{(i,j-1)}+b \\
%	V_{(i,j)}+\delta(S_i,S_j) \\
%	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)}
%\end{array}\right.\\\\
%V_{(i,j)}&=&\min\left\{\begin{array}{ll}
%	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
%	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
%	V_{(i+1,j-1)}+es(i,j) \\
%	VBI(i,j) \\
%	\min_{i<k<j-1}\{W_{(i+,k)}+W_{(k+1,j-1)}\} +c
%\end{array}\right.\\\\
% F_{(j)}&=&\min\left\{\begin{array}{ll}
%F_{(j-i)} \\
%\min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
%\end{array}\right. \text{(Free Energy)}\\\\
%VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
%\end{array}\]
%
%XXX: missing: eh(i,j), ebi(i,j,i',j'), es(i,j), ...

>>>>

% ----------------------------------------------
\subsection{Convex polygon triangulation}
Problem: triangulating a polygon with least cost overall for added edges

Alphabets:\ul
\item Input: $\Sigma(S)= $ a matrix of costs for every possible edge in the polygon. The 
size of the input is $\frac{n^2}{2}$, where $n$ is the number of vertices in the polygon. 
\item Cost matrix: $\Sigma(M)=\mathbb{N}$. The sum of cost of edges is not "bounded".
\item Backtrack matrix: $\Sigma(B)=\{n\}$
\ule

Recurrence:
\[M_{(i,j)}= \max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} + c(i,k) | k = B_{(i,j)} \]

It is interesting to note that even in the sequential world, this problem is best solved 
by filling the diagonals, ie. computing sub-solutions for all polygons on size $k$ before
those of size $k+1$.

% ----------------------------------------------

@@

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	M_{(i+1,j-1)}+\omega(i,j) & SW\\
	M_{(i+1,j)} & S\\
	M_{(i,j-1)} & W\\
	\max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} & k
\end{array}\right\} = B_{(i,j)} \]
With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Zuker folding}
XXX: we let this aside for the moment as the recurrence has too many unknown parameters
Problem: folding a RNA string $S$ over itself.

Alphabets:\ul
\item Input: $\Sigma(S)=\{A,C,G,U\}$.
\item Cost matrix: $\Sigma(W)=\Sigma(V)=\Sigma(F)=\{0..n\},n=$
\item Backtrack matrix: $\Sigma(B)=\{\}$
\ule

Initialization:\ul
\item Cost matrices:\ul
	\item xxx
\ule
\item Backtrack matrix: $B$
\ule

Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l}
	W_{(i+1,j)}+b \\
	W_{(i,j-1)}+b \\
	V_{(i,j)}+\delta(S_i,S_j) \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)}
\end{array}\right.\\\\

V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+1,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\\\

VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c \\\\

F_{(j)}&=&\min\left\{\begin{array}{ll}
F_{(j-i)} \\
\min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
\end{array}\right. \text{(Free Energy)}\\\\

\end{array}\]

\subsection{Comments}

VBI stores scores for internal loops and bulges.
Notice in VBI, two moving indices $i', j'$. In general this gives a quadratic complexity for that particular recurrence, making the overall complexity of the algorithm $O(n^4)$. In practice (as read in Arpith's work), we have that $i' - i + j'- j < 30$, so the complexity is reduced to $O(n^3)$, albeit with a large constant.

%XXX: missing: eh(i,j), ebi(i,j,i',j'), es(i,j), ...
